---
title: "Practical Machine Learning - Peer Assessment"
author: "Chris Chu"
date: "24 May, 2015"
output: html_document
---

#Summary
*The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. Thanks for their generosity in allowing their data to be used for my assignment. 
*This analysis predicts the subjects' manner in performing weight lifting exercises. This analysis collects the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. The analysis will use prediction model to predict 20 different test cases. 

#Import Data
* load the necessary library.
* load the data.
```{r}
library(caret)
library(randomForest)

testing         <- read.table("pml-testing.csv" ,sep = "," , header = TRUE, na.strings = c("", "NA", "NULL"))
training        <- read.table("pml-training.csv" , sep = "," , header = TRUE, na.strings = c("", "NA", "NULL"))
```

#Data checking
* Check whether the dimension of of the training and testing dataset match.
* Check whether the dataset has N/A values
```{r}
dim(training)
dim(testing)

trainingna      <- sapply(training, function(x) {sum(is.na(x))})
table(trainingna)
testingna       <- sapply(testing, function(x) {sum(is.na(x))})
table(testingna)
```

#Data extraction
* Extract the row without N/A values into our new dataset
* Remove all personal and time columns
```{r}
training        <- training[, colSums(is.na(training)) == 0]
testing         <- testing[, colSums(is.na(testing)) == 0]

training        <- training[, c(7:60)]
testing         <- testing[, c(7:60)]
```

#Data spliting
* Split the training data into training and testing datasets. 
* The splitting is 60% for the training and 40% for the testing.
```{r}
inTrain         <- createDataPartition(y = training$classe, p=0.60, list=FALSE)
trainmodel      <- training[inTrain,]
testmodel       <- training[-inTrain,]
```

#Modeling
* Use the random forests machine learning method with the 60% dataset
```{r}
model <- randomForest(classe ~ . , data = trainmodel)

model
```

#Out of Sample Accuracy
* Calculate the out of sample accuracy with the 40% dataset
```{r}
testmodel_prediction <- predict(model, testmodel)
confusionMatrix(testmodel_prediction, testmodel$classe)
```
* From the table, the model shows 99.6% accuracy.

#I do not have enought time to fninsh the prediction this week. I will continue to work and hand in next week.